<h1>Remembering: the victims of the Troubles</h1>
<p>All data is extracted from Michael McKeown's open dataset (available at https://cain.ulster.ac.uk/victims/mckeown/index.html)</p>
```{r}
library(tidyverse)
library(ggplot2)
library(janitor)
library(extrafont)
library(RColorBrewer)
mckeown <- read.csv('deaths_mckeown.csv', na =c('', NA))
glimpse(mckeown)

mc_clean <- mckeown %>%
  clean_names() #%>%
 # mutate(year = parse_date(year, format = '%Y'))

 mc_clean %>%
  group_by(year) %>%
  summarise(victims = n()) -> victims_per_year
  # write.csv('deaths_by_year.csv', row.names = FALSE)
```
<h2>Victims per year</h2>
<p>From the 1970s peak to gradual decline, but tension remains high</p>
```{r}
ggplot(victims_per_year, aes(year, victims)) + geom_area(col = '#801515', fill = '#D46A6A', alpha = .5) +
  theme_minimal() +
  labs(title = 'Victims in NI conflict, 1969-2001', caption = 'Dati: McKeown, Michael. (2009). Spreadsheet of Deaths Associated with Violence in Northern Ireland, 1969-2001', x = '', y = 'Vittime') +
  theme(text = element_text(family = 'Open Sans'), axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)), plot.title = element_text(margin = margin(t = 0, r = 0, b = 20, l = 0)), panel.grid = element_line(colour = '#f2f2f2'), panel.grid.minor.x = element_blank())
```
<h2>Victims by rationale</h2>
```{r}
vic_by_rationale <- mc_clean %>%
  group_by(rationale) %>%
  summarise(victims = n()) %>%
  arrange(desc(victims))

write.csv(vic_by_rationale, 'vic_rationale.csv', row.names = FALSE)
ggplot(vic_by_rationale, aes(reorder(rationale, victims), victims)) +
  geom_col(alpha = .7, fill = '#801515') +
  coord_flip() + 
  theme_minimal() +
  labs(x = '', y = 'Victims', title = 'Victims in NI conflict by rationale, 1969-2001', caption = 'Data: McKeown, Michael. (2009). Spreadsheet of Deaths Associated with Violence in Northern Ireland, 1969-2001') +
  theme(text = element_text(family = 'Open Sans'), plot.title = element_text(margin = margin(b = 15)), axis.title.x = element_text(margin = margin(t = 10, b = 10)), panel.grid.major.y = element_blank(), panel.grid = element_line(colour = '#f2f2f2'))
```
<h2>Victims by religion</h2>
<p>Overview</p>
```{r}
vic_by_religion <- mc_clean %>%
  group_by(religion) %>%
  summarise(victims = n()) %>%
  arrange(desc(victims))

ggplot(vic_by_religion, aes(reorder(religion, victims), victims, fill = religion)) +
  geom_col(alpha = .7, position = 'stack' ) +
  coord_flip() + 
  theme_minimal() +
  labs(x = '', y = 'Victims', title = 'Victims in NI conflict by religion, 1969-2001', caption = 'Data: McKeown, Michael. (2009). Spreadsheet of Deaths Associated with Violence in Northern Ireland, 1969-2001') +
  theme(text = element_text(family = 'Open Sans'), plot.title = element_text(margin = margin(b = 15)), axis.title.x = element_text(margin = margin(t = 10, b = 10)), panel.grid.major.y = element_blank(), panel.grid = element_line(colour = '#f2f2f2'))+
  scale_fill_manual(values = c('#80CDC1', '#CCCCCC', '#E6AB02'), guide = FALSE )
```

```{r}
# qualitative palette for the next two charts
qual_p <- brewer.pal(n = 4, name = 'Set2')
```
<p>Agency-group match is done based on McKeon's lookup tables</p>

<h3>Victims identified as catholic</h3>
```{r}
catholic_victims <- mc_clean %>%
  filter(religion == 'Catholic')

lookup_agency <- read.csv('lookup_agency.csv')

indices <- match ( x = catholic_victims$agency, table = lookup_agency$Agency)

lookup <- add_column(lookup_agency[indices, ], x = catholic_victims$agency)

catholic_victims <- catholic_victims %>%
  mutate('group' = lookup$Group) %>%
  group_by(group) %>%
  summarise(victims = n())

ggplot(catholic_victims, aes(reorder(group, victims), victims, group = group, fill = group)) + geom_col(alpha = .7, guides= FALSE) +
  coord_flip() +
  theme_minimal() +
labs(x = '', y = 'Victims', title = 'Catholic victims in NI conflict by group, 1969-2001', caption = 'Data: McKeown, Michael. (2009). Spreadsheet of Deaths Associated with Violence in Northern Ireland, 1969-2001') +
  theme(text = element_text(family = 'Open Sans'), plot.title = element_text(margin = margin(b = 15)), axis.title.x = element_text(margin = margin(t = 10, b = 10)), panel.grid.major.y = element_blank(), panel.grid = element_line(colour = '#f2f2f2')) +
  scale_fill_manual(values = qual_p, guide = FALSE)
```
<h3>Victims identified as protestant</h3>
```{r}
protestant_victims <- mc_clean %>%
  filter(religion == 'Protestant')



indices_p <- match ( x = protestant_victims$agency, table = lookup_agency$Agency)

lookup_p <- add_column(lookup_agency[indices_p, ], x = protestant_victims$agency)

protestant_victims <- protestant_victims %>%
  mutate('group' = lookup_p$Group) %>%
  group_by(group, agency) %>%
  summarise(victims = n())

ggplot(protestant_victims, aes(reorder(group, victims), victims, group = group, fill = group)) + 
  geom_col(alpha = .7) +
  coord_flip() +
  theme_minimal() +
labs(x = '', y = 'Victims', title = 'Protestant victims in NI conflict by group, 1969-2001', caption = 'Data: McKeown, Michael. (2009). Spreadsheet of Deaths Associated with Violence in Northern Ireland, 1969-2001') +
  theme(text = element_text(family = 'Open Sans'), plot.title = element_text(margin = margin(b = 15)), axis.title.x = element_text(margin = margin(t = 10, b = 10)), panel.grid.major.y = element_blank(), panel.grid = element_line(colour = '#f2f2f2')) +
  scale_fill_manual(values = qual_p, guide = FALSE)
```

```{r}
names(lookup_agency)[2] <- 'agency'


mc_clean %>%
  left_join(lookup_agency, by = 'agency') -> mc_clean

mc_clean %>%
  group_by(religion, Group, agency) %>%
  count(religion) %>%
  ggplot(aes(religion, n, fill = Group)) + geom_col(position = 'fill') + coord_flip() +
  theme_minimal() +
  theme(panel.grid = element_blank(), legend.position = 'top')
```

```{r}
mc_clean %>%
  filter(multiple_fatality == TRUE) %>%
  group_by(year, location_name, x1st_fatality, context, year) %>%
  count(x1st_fatality) %>%
  arrange(desc(n)) %>%
  ungroup() %>%
  top_n(10) %>%
 write.csv('top15_fatality.csv', row.names = FALSE)
  ggplot(aes(reorder(x1st_fatality, n), n, fill = context)) + geom_col() + coord_flip() +
  geom_text(aes(x1st_fatality, n - 3, label = year)) +
  theme_minimal()
```

```{r}
mc_clean %>%
  group_by(sex, status) %>%
  count(status) %>% 
  ggplot(aes(sex, n, fill = status)) + geom_col(position = 'stack')
```

```{r}
mc_clean %>%
  group_by(location_name) %>%
  summarise(victims = n()) %>%
  arrange(desc(victims)) %>%
  write.csv('victims_location.csv', row.names = FALSE)
```

```{r}
lookup.status <- read.csv('lookup_status.csv') %>%
  clean_names()



mc_clean %>%
  left_join(., lookup.status, by = 'status') %>%
  group_by(religion, group, status) %>%
  summarise(victims= n()) -> status_groups

status_groups %>%
  mutate(child = paste(group,status, sep = ' - ')) -> status_groups

write.csv(status_groups, 'merged.csv', row.names = FALSE)

# religions <- tibble(name = unique(status_groups$religion)) %>%
#   mutate(parent = rep('Religion', nrow(.)), victims = rep(NA, nrow(.))) %>%
#   write.csv('religions.csv', row.names = FALSE)
# 
# groups <- unique(status_groups[,c('religion','group')])
# names(groups) <- c('parent', 'name')
# groups$victims <- NA
# write.csv(groups, 'groups.csv', row.names = FALSE)
# statuses <- (status_groups[,c('group','status', 'victims')])
# names(statuses) <- c('parent', 'name', 'victims')
# 
# write.csv(statuses, 'statuses.csv', row.names = FALSE)
# 
# 
# merge(religions, groups) -> df
# 
# 
# distinct(status_groups$religion, status_groups$group)
# status_groups %>%
#   gather(key = name, value = n, - victims)
# 
# for (i in 1:length(status_groups)) {
#   if ()
# },
```



